### 2018-01-26 manuscript.rmd
added details about Rudjakovia & SRA numbers

Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea. Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators. A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora. Here, we present a new siphonophore phylogeny based on new transcriptome data from 30 siphonophore species analyzed in combination with 13 publicly available genomic and transcriptomic datasets. We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including sexual system (monoecy vs. dioecy), gain and loss of zooid types, life history traits, and habitat. The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but we find strong support for new clades within Codonophora that were previously unresolved. These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose twice.

@@

xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens collected for this work. New data indicated by Y, blank fields indicate that data were already published.")
print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.75')

@@

Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig. 5A), once in Calycophorae and once in Clade A. This is consistent with differences in the arrangements of male and female gonophores in the two clades. In Clade A, male and female zooids are found within the same cormidium (a single reiterated sequence of zooids along the stem, see Fig. 2). In these species, the male and female zooids are placed at very different but well defined locations within the cormidium. Meanwhile in calycophorans, each cormidium bears either male or female gonophores. In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row. In either case, male and female zooids are found at the same corresponding locations within the cormidia. In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious. Both Calycophorae and Clade A have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode. Similar independent transitions from gonochorism (separate sex) to hermaphroditism (both sexes in the same individual) have been identified in shallow-water scleractinian corals (Anthozoa, Cnidaria) [@kerr2011correlated]. To test this hypothesis, a more extensive taxon sampling of the Calycophorae is needed.

@@

![Stochastic mapping reconstruction of the evolutionary history of A) sexual mode, whether a colony is monoecious or dioecious and B) presence/absence of palpons (modified reduced gastrozooids). The color gradients show the reconstructed probability estimate of the discrete character states along the branches. Intermediate values reflect uncertainty.](figures/Figure5.pdf)

@@

Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. Among other relationships, we identify the clade Euphysonectae as the sister group to Calycophorae. Our results suggest that monoecy arose twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.  We are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora, which will require greater taxon sampling and improved morphological understanding of many poorly known species. The improved resolution presented in this study suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae (where we sampled 13 of 62 valid described species that likely belong to the group) and Calycophorae (where we sampled 9 species in a clade of 109 valid described species) to further resolve the internal relationships within these clades.

>>>>

Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea. Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators. A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora. Here, we present a new siphonophore phylogeny based on new transcriptome data from 30 siphonophore species analyzed in combination with 13 publicly available genomic and transcriptomic datasets. We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including sexual system (monoecy vs. dioecy), gain and loss of zooid types, life history traits, and habitat. The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but we find strong support for new clades within Codonophora that were previously unresolved. These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose at least twice.

@@

specimen_data_reduced<-as.data.frame(specimen_data_reduced)
xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens collected for this work. New data indicated by Y, blank fields indicate that data were already published.")
print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.7')

@@

Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig. 5A), once in Calycophorae and once in Clade A. This is consistent with differences in the arrangements of male and female gonophores in the two clades. In Clade A, male and female zooids are found within the same cormidium (a single reiterated sequence of zooids along the stem, see Fig. 2). In these species, the male and female zooids are placed at very different but well defined locations within the cormidium. Meanwhile in calycophorans, each cormidium bears either male or female gonophores. In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row. In either case, male and female zooids are found at the same corresponding locations within the cormidia. In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious. The sexual system and cormidial arrangement of *Rudjakovia* is undescribed, although preliminary observations suggest that this species may be monoecious and that monoecy arose a third time in the Pyrostephidae. A detailed redescription of *Rudjakovia* would help clarify this.

Both Calycophorae and Clade A have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode. Similar independent transitions from gonochorism (separate sex) to hermaphroditism (both sexes in the same individual) have been identified in shallow-water scleractinian corals (Anthozoa, Cnidaria) [@kerr2011correlated]. To test this hypothesis, a more extensive taxon sampling of the Calycophorae is needed.

@@

![Stochastic mapping reconstruction of the evolutionary history of A) sexual mode, whether a colony is monoecious or dioecious and B) presence/absence of palpons (modified reduced gastrozooids). The color gradients show the reconstructed probability estimate of the discrete character states along the branches. Intermediate values reflect uncertainty. The grey dashed branch leading to *Rudjakovia* sp. indicates that the sexual mode of this species is unknown.](figures/Figure5.pdf)

@@

Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support. Among other relationships, we identify the clade Euphysonectae as the sister group to Calycophorae. Our results suggest that monoecy arose at least twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.  We are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora, which will require greater taxon sampling and improved morphological understanding of many poorly known species. The improved resolution presented in this study suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae (where we sampled 13 of 62 valid described species that likely belong to the group) and Calycophorae (where we sampled 9 species in a clade of 109 valid described species) to further resolve the internal relationships within these clades.
